Supplementary Information

### 2021-01-31 doc/manuscript/MEE_manuscript.Rmd
add recruitment plots
#109

# SUPPLEMENTARY INFO (this chunk must be repeated in MEE_manuscript_SI.Rmd)

#Appendices
site_details_appendix <- "S1"
DBH_reconstruction <- "S2"
climate_correction <- "S3"
traditional_comparison <- "S4"
changing_climate <- "S5"

@@

#SI figure numbers
all_traditional_comparisons = "S1-S4"
SCBI_traditional_comparison = "S1"
Zofin_traditional_comparison = "S2"
CB_traditional_comparison = "S3"
SC_traditional_comparison = "S4"

all_climwin = "S5-S7"
climwin_sameV_sameW = "S5"
climwin_sameV_diffW = "S6"
climwin_diffV_diffW = "S8"

all_GLS = "S9-S28"
GLS_BCNM = "S9"
GLS_BCNM_year = "S10"
GLS_HKK = "S11"
GLS_HKK_year = "S12"
GLS_SCBI = "S13"
GLS_SCBI_year = "S14"
GLS_LDW = "S15"
GLS_LDW_year = "S16"
GLS_HF = "S17"
GLS_HF_year = "S18"
GLS_ZOF = "S19"
GLS_ZOF_year = "S20"
GLS_NIO = "S21"
GLS_NIO_year = "S22"
GLS_LT = "S23"
GLS_LT_year = "S24"
GLS_CB = "S25"
GLS_CB_year = "S26"
GLS_SC = "S27"
GLS_SC_year = "S28"

SC_1970break = "S29"

RW_interactions_all= "S30"
BAI_interactions_all= "S31"

decadal_analysis_figs_all = "S32-S41"
decadal_plot_BCNM = "S32"
decadal_plot_HKK = "S33"
decadal_plot_SCBI = "S34"
decadal_plot_LDW = "S35"
decadal_plot_HF = "S36"
decadal_plot_ZOF = "S37"
decadal_plot_NIO = "S38"
decadal_plot_LT = "S39"
decadal_plot_CB = "S40"
decadal_plot_SC = "S41"

#SI figure captions
SCBI_traditional_comparison_legend <- "Figure S1. Comparison of our approach with traditional methods of identifying climate signals: LITU at SCBI."
Zofin_traditional_comparison_legend <- "Figure S2. Comparison of our approach with traditional methods of identifying climate signals: ABAL at Zofin."
CB_traditional_comparison_legend <- "Figure S3. Comparison of our approach with traditional methods of identifying climate signals: PSME at Cedar Breaks."
SC_traditional_comparison_legend <- "Figure S4. Comparison of our approach with traditional methods of identifying climate signals: PIMA at Scotty Creek."

climwin_sameV_sameW_legend = "Figure S5. (PRE at SCBI)"
climwin_sameV_diffW_legend = "Figure S6. (PET at SCBI)"
climwin_diffV_diffW_legend = "Figure S7. (TMX/TMP at HKK)"

# without year
GLS_BCNM_legend = "Figure S9. Best GLS models including climate and DBH for Barro Colorado Nature Monument (Panama)"
GLS_BCNM_year_legend = "Figure S10. Best GLS models including climate, DBH, and year for Barro Colorado Nature Monument (Panama)"
GLS_HKK_legend = "Figure S11. Best GLS models including climate and DBH for Huai Kha Khaeng (Thailand)"
GLS_HKK_year_legend = "Figure S12. Best GLS models including climate, DBH, and year for Huai Kha Khaeng (Thailand)"
GLS_SCBI_legend = "Figure S13. Best GLS models including climate and DBH for the Smithsonian Conservation Biology Institute (Virginia, USA)"
GLS_SCBI_year_legend = "Figure S14. Best GLS models including climate, DBH, and year for the Smithsonian Conservation Biology Institute (Virginia, USA)"
GLS_LDW_legend = "Figure S15. Best GLS models including climate and DBH for Lilley Dickey Woods (Indiana, USA)"
GLS_LDW_year_legend = "Figure S16. Best GLS models including climate, DBH, and year for Lilley Dickey Woods (Indiana, USA)"
GLS_HF_legend = "Figure S17. Best GLS models including climate and DBH for Harvard Forest (Massachusetts, USA) "
GLS_HF_year_legend = "Figure S18. Best GLS models including climate, DBH, and year for Harvard Forest (Massachusetts, USA) "
GLS_ZOF_legend = "Figure S19. Best GLS models including climate and DBH for Zofin Forest (Czech Republic)"
GLS_ZOF_year_legend = "Figure S20. Best GLS models including climate, DBH, and year for Zofin Forest (Czech Republic)"
GLS_NIO_legend = "Figure S21. Best GLS models including climate and DBH for Niobrara (Nebraska, USA)"
GLS_NIO_year_legend = "Figure S22. Best GLS models including climate, DBH, and year for Niobrara (Nebraska, USA)"
GLS_LT_legend = "Figure S23. Best GLS models including climate and DBH for Little Tesuque (New Mexico, USA)"
GLS_LT_year_legend = "Figure S24. Best GLS models including climate, DBH, and year for Little Tesuque (New Mexico, USA)"
GLS_CB_legend = "Figure S25. Best GLS models including climate and DBH for Cedar Breaks (Utah, USA)"
GLS_CB_year_legend = "Figure S26. Best GLS models including climate, DBH, and year for Cedar Breaks (Utah, USA)"
GLS_SC_legend = "Figure S27. Best GLS models including climate and DBH for Scotty Creek (Northwest Territory, Canada)"
GLS_SC_year_legend = "Figure S28. Best GLS models including climate, DBH, and year for Scotty Creek (Northwest Territory, Canada)"

SC_1970break_legend = "Figure S29. Climate responses at Scotty Creek (Northwest Territory, Canada) before and after 1970."

RW_interactions_all_legend = "Figure S30. (RW_interactions_all)"
BAI_interactions_all_legend ="Figure S31. (BAI_interactions_all)"

decadal_plot_BCNM_legend = "Figure S32. (Decadal BCNM)"
decadal_plot_HKK_legend = "Figure S33. (Decadal HKK)"
decadal_plot_SCBI_legend = "Figure S34. (Decadal SCBI)"
decadal_plot_LDW_legend = "Figure S35. (Decadal LDW)"
decadal_plot_HF_legend = "Figure S36. (Decadal HF)"
decadal_plot_ZOF_legend = "Figure S37. (Decadal ZOF)"
decadal_plot_NIO_legend = "Figure S38. (Decadal NIO)"
decadal_plot_LT_legend = "Figure S39. (Decadal LT)"
decadal_plot_CB_legend = "Figure S40. (Decadal CB)"
decadal_plot_SC_legend = "Figure S41. (Decadal SC)"

@@

We analyzed previously collected tree-ring data from ten sites ranging from 9.15$^\circ$ to 61.30$^\circ$ N latitude and representing a wide range of forest and tree types: tropical broadleaf deciduous and evergreen, temperate broadleaf deciduous and needleleaf evergreen, and boreal needleleaf evergreen (Tables `r sites`, `r site_details`, `r species_table`).
Nine of these sites (exception: LT) are large forest dynamics plots of the Forest Global Earth Observatory [ForestGEO; @anderson-teixeira_ctfs-forestgeo_2015; @davies_forestgeo_2021].
Trees of species represented within the ForestGEO plots were cored within the plot (n=#) and/or within 10 km (n=#), following a variety of sampling protocols designed to meet the varied objectives of the original studies (Tables `r site_details`, `r sampling_details_table`). 
In using this diversity of data sources, we ensured that our approach was able to handle challenges presented by varying methodologies and forest types.

@@

The majority of negative growth trends observed here (Fig. `r year_fig`) are probably attributable to successional stand dynamics as cohorts and stands age. 
For species exhibiting a pulse of recruitment in the past followed by little subsequent recruitment (e.g., **SPECIES, FIGURES**), persistent differences in growth rates among individuals would naturally lead to a trend of declining growth, as faster-growing individuals reach various size thresholds earlier [@brienen_tree_2017].
Such a trend is not indicative of a directional response to changing environmental conditions that would be generalizable across stands (e.g., rising atmospheric CO~2~), but it does indicate declining growth of the species within the stand. 
This can occur, for example, when a light-demanding species is no longer recruiting under a closed canopy (e.g., *Quercus* spp. at SCBI), such that fast-growing individuals disappear from progressively larger cohorts as time proceeds. 
In secondary stands, particularly those where the majority of trees established at a similar time (**SCBI, HF,...**; Appendix `r site_details_appendix` **FIGURES**), growth declines are consistent with the tendency for faster tree growth during early succession (**REFS**), and with increasing competition and declining woody productivity as young stands mature [e.g., @pregitzer_carbon_2004; @goulden_patterns_2011].
Even within mature forests, light-demanding species that establish in gaps (e.g., *Jacaranda copaia* at BCNM, *Populus tremuloides* at CB; Table `r species_table`) would tend to experience an increasingly competitive environment through time.
For more shade-tolerant species in stands with no known disturbance within the past 1.5 centuries (**ZOF?, NM?, CB, SC**), trends would represent more gradual successional changes or responses to global change.
At Cedar Breaks, which has tree-ring records dating back to 0 AD (Cooper & Lutz, unpublished data), mixed growth trends probably reflect later successional changes in a mixed-disturbance forest and/or shifting competitive advantages, perhaps in part driven by changing environmental conditions.

>>>>

# SUPPLEMENTARY INFO (this chunk must be repeated in MEE_manuscript_SI.Rmd)
#Appendices
site_details_appendix <- "S1"
DBH_reconstruction <- "S2"
climate_correction <- "S3"
traditional_comparison <- "S4"
changing_climate <- "S5"

@@

#SI figure numbers
all_recruitment_plots="S1-S10"

all_traditional_comparisons = "S11-S14"
SCBI_traditional_comparison = "S11"
Zofin_traditional_comparison = "S12"
CB_traditional_comparison = "S13"
SC_traditional_comparison = "S14"

all_climwin = "S15-S17"
climwin_sameV_sameW = "S15"
climwin_sameV_diffW = "S16"
climwin_diffV_diffW = "S17"

all_GLS = "S18-S37"
GLS_BCNM = "S18"
GLS_BCNM_year = "S19"
GLS_HKK = "S20"
GLS_HKK_year = "S21"
GLS_SCBI = "S22"
GLS_SCBI_year = "S23"
GLS_LDW = "S24"
GLS_LDW_year = "S25"
GLS_HF = "S36"
GLS_HF_year = "S27"
GLS_ZOF = "S28"
GLS_ZOF_year = "S29"
GLS_NIO = "S30"
GLS_NIO_year = "S31"
GLS_LT = "S32"
GLS_LT_year = "S33"
GLS_CB = "S34"
GLS_CB_year = "S35"
GLS_SC = "S36"
GLS_SC_year = "S37"

SC_1970break = "S38"

RW_interactions_all= "S39"
BAI_interactions_all= "S40"

decadal_analysis_figs_all = "S41-S50"
decadal_plot_BCNM = "S41"
decadal_plot_HKK = "S42"
decadal_plot_SCBI = "S43"
decadal_plot_LDW = "S44"
decadal_plot_HF = "S45"
decadal_plot_ZOF = "S46"
decadal_plot_NIO = "S47"
decadal_plot_LT = "S48"
decadal_plot_CB = "S49"
decadal_plot_SC = "S50"

#SI figure captions
recruitment_plot_BCNM_legend = "Figure S1. (recruitment BCNM)"
recruitment_plot_HKK_legend = "Figure S2. (recruitment HKK)"
recruitment_plot_SCBI_legend = "Figure S3. (recruitment SCBI)"
recruitment_plot_LDW_legend = "Figure S4. (recruitment LDW)"
recruitment_plot_HF_legend = "Figure S5. (recruitment HF)"
recruitment_plot_ZOF_legend = "Figure S6. (recruitment ZOF)"
recruitment_plot_NIO_legend = "Figure S7. (recruitment NIO)"
recruitment_plot_LT_legend = "Figure S8. (recruitment LT)"
recruitment_plot_CB_legend = "Figure S9. (recruitment CB)"
recruitment_plot_SC_legend = "Figure S10. (recruitment SC)"

SCBI_traditional_comparison_legend <- "Figure S11. Comparison of our approach with traditional methods of identifying climate signals: LITU at SCBI."
Zofin_traditional_comparison_legend <- "Figure S12. Comparison of our approach with traditional methods of identifying climate signals: ABAL at Zofin."
CB_traditional_comparison_legend <- "Figure S13. Comparison of our approach with traditional methods of identifying climate signals: PSME at Cedar Breaks."
SC_traditional_comparison_legend <- "Figure S14. Comparison of our approach with traditional methods of identifying climate signals: PIMA at Scotty Creek."

climwin_sameV_sameW_legend = "Figure S15. (PRE at SCBI)"
climwin_sameV_diffW_legend = "Figure S16. (PET at SCBI)"
climwin_diffV_diffW_legend = "Figure S17. (TMX/TMP at HKK)"

GLS_BCNM_legend = "Figure S18. Best GLS models including climate and DBH for Barro Colorado Nature Monument (Panama)"
GLS_BCNM_year_legend = "Figure S19. Best GLS models including climate, DBH, and year for Barro Colorado Nature Monument (Panama)"
GLS_HKK_legend = "Figure S20. Best GLS models including climate and DBH for Huai Kha Khaeng (Thailand)"
GLS_HKK_year_legend = "Figure S21. Best GLS models including climate, DBH, and year for Huai Kha Khaeng (Thailand)"
GLS_SCBI_legend = "Figure S22. Best GLS models including climate and DBH for the Smithsonian Conservation Biology Institute (Virginia, USA)"
GLS_SCBI_year_legend = "Figure S23. Best GLS models including climate, DBH, and year for the Smithsonian Conservation Biology Institute (Virginia, USA)"
GLS_LDW_legend = "Figure S24. Best GLS models including climate and DBH for Lilley Dickey Woods (Indiana, USA)"
GLS_LDW_year_legend = "Figure S25. Best GLS models including climate, DBH, and year for Lilley Dickey Woods (Indiana, USA)"
GLS_HF_legend = "Figure S26. Best GLS models including climate and DBH for Harvard Forest (Massachusetts, USA) "
GLS_HF_year_legend = "Figure S27. Best GLS models including climate, DBH, and year for Harvard Forest (Massachusetts, USA) "
GLS_ZOF_legend = "Figure S28. Best GLS models including climate and DBH for Zofin Forest (Czech Republic)"
GLS_ZOF_year_legend = "Figure S29. Best GLS models including climate, DBH, and year for Zofin Forest (Czech Republic)"
GLS_NIO_legend = "Figure S30. Best GLS models including climate and DBH for Niobrara (Nebraska, USA)"
GLS_NIO_year_legend = "Figure S31. Best GLS models including climate, DBH, and year for Niobrara (Nebraska, USA)"
GLS_LT_legend = "Figure S32. Best GLS models including climate and DBH for Little Tesuque (New Mexico, USA)"
GLS_LT_year_legend = "Figure S33. Best GLS models including climate, DBH, and year for Little Tesuque (New Mexico, USA)"
GLS_CB_legend = "Figure S34. Best GLS models including climate and DBH for Cedar Breaks (Utah, USA)"
GLS_CB_year_legend = "Figure S35. Best GLS models including climate, DBH, and year for Cedar Breaks (Utah, USA)"
GLS_SC_legend = "Figure S36. Best GLS models including climate and DBH for Scotty Creek (Northwest Territory, Canada)"
GLS_SC_year_legend = "Figure S37. Best GLS models including climate, DBH, and year for Scotty Creek (Northwest Territory, Canada)"

SC_1970break_legend = "Figure S38. Climate responses at Scotty Creek (Northwest Territory, Canada) before and after 1970."

RW_interactions_all_legend = "Figure S39. (RW_interactions_all)"
BAI_interactions_all_legend ="Figure S40. (BAI_interactions_all)"

decadal_plot_BCNM_legend = "Figure S41. (Decadal BCNM)"
decadal_plot_HKK_legend = "Figure S42. (Decadal HKK)"
decadal_plot_SCBI_legend = "Figure S43. (Decadal SCBI)"
decadal_plot_LDW_legend = "Figure S44. (Decadal LDW)"
decadal_plot_HF_legend = "Figure S45. (Decadal HF)"
decadal_plot_ZOF_legend = "Figure S46. (Decadal ZOF)"
decadal_plot_NIO_legend = "Figure S47. (Decadal NIO)"
decadal_plot_LT_legend = "Figure S48. (Decadal LT)"
decadal_plot_CB_legend = "Figure S49. (Decadal CB)"
decadal_plot_SC_legend = "Figure S50. (Decadal SC)"

@@

We analyzed previously collected tree-ring data from ten sites ranging from 9.15$^\circ$ to 61.30$^\circ$ N latitude and representing a wide range of forest and tree types: tropical broadleaf deciduous and evergreen, temperate broadleaf deciduous and needleleaf evergreen, and boreal needleleaf evergreen (Tables `r sites`, `r site_details`, `r species_table`).
Nine of these sites (exception: LT) are large forest dynamics plots of the Forest Global Earth Observatory [ForestGEO; @anderson-teixeira_ctfs-forestgeo_2015; @davies_forestgeo_2021].
Trees of species represented within the ForestGEO plots were cored within the plot (n=#) and/or within 10 km (n=#), following a variety of sampling protocols designed to meet the varied objectives of the original studies (Tables `r site_details`, `r sampling_details_table`). 
There was wide variation in the distribution of record start years, which reflect--with some error--the age distribution of trees cored (Figs. `r all_recruitment_plots`).
In using this diversity of data sources, we ensured that our approach was able to handle challenges presented by varying methodologies and forest types.

@@

The majority of negative growth trends observed here (Fig. `r year_fig`) are probably attributable to successional stand dynamics as cohorts and stands age. 
For species exhibiting a pulse of recruitment in the past followed by little subsequent recruitment (e.g., *Acer rubrum* and *Betula alleghaniensis* at HF; Figs. `r all_recruitment_plots`), persistent differences in growth rates among individuals would naturally lead to a trend of declining growth, as faster-growing individuals reach various size thresholds earlier [@brienen_tree_2017].
Such a trend is not indicative of a directional response to changing environmental conditions that would be generalizable across stands (e.g., rising atmospheric CO~2~), but it does indicate declining growth of the species within the stand. 
This can occur, for example, when a light-demanding species is no longer recruiting under a closed canopy (e.g., *Quercus* spp. at SCBI), such that fast-growing individuals disappear from progressively larger $DBH$ classes as time proceeds. 
In secondary stands, particularly those where the majority of trees established at a similar time (**SCBI, HF,...**; Appendix `r site_details_appendix`, Figs. `r all_recruitment_plots`), growth declines are consistent with the tendency for faster tree growth during early succession (**REFS**), and with increasing competition and declining woody productivity as young stands mature [e.g., @pregitzer_carbon_2004; @goulden_patterns_2011].
Even within older forests, light-demanding species that establish in gaps (e.g., *Jacaranda copaia* at BCNM, *Populus tremuloides* at CB; Table `r species_table`) would tend to experience an increasingly competitive environment through time.
For more shade-tolerant species in stands with no known disturbance within the past 1.5 centuries (**ZOF?, NM?, CB, SC**), trends would represent more gradual successional changes or responses to global change.
At Cedar Breaks, which has tree-ring records dating back to 0 AD (Cooper & Lutz, unpublished data), mixed growth trends probably reflect later successional changes in a mixed-disturbance forest and/or shifting competitive advantages, perhaps in part driven by changing environmental conditions.

### 2021-01-31 doc/manuscript/MEE_manuscript_SI.Rmd
add recruitment plots
#109

@@

@@

![`r decadal_plot_CB_legend`](tables_figures/SI_figures/decadal_plots/CB.png){ width=100% }
\newpage
## `r decadal_plot_SC_legend`

>>>>

#recruitment plots
do.call(file.remove, list(list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/doc/manuscript/tables_figures/SI_figures/recruitment_years/", full.names = TRUE))) # first remove old files
flist <- list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO_dendro/data_processed/", "^Recrutement_years_.+[.]png$",  full.names = TRUE)
file.copy(flist, "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/doc/manuscript/tables_figures/SI_figures/recruitment_years/")

@@

## `r recruitment_plot_BCNM_legend`

![`r recruitment_plot_BCNM_legend`](tables_figures/SI_figures/recruitment_years/Recrutement_years_BCI.png){ width=50% }

\newpage
## `r recruitment_plot_HKK_legend`

![`r recruitment_plot_HKK_legend`](tables_figures/SI_figures/recruitment_years/Recrutement_years_HKK.png){ width=50% }

\newpage
## `r recruitment_plot_SCBI_legend`

![`r recruitment_plot_SCBI_legend`](tables_figures/SI_figures/recruitment_years/Recrutement_years_SCBI.png){ width=50% }
\newpage

## `r recruitment_plot_LDW_legend`

![`r recruitment_plot_LDW_legend`](tables_figures/SI_figures/recruitment_years/Recrutement_years_LillyDickey.png){ width=50% }

\newpage
## `r recruitment_plot_HF_legend`

![`r recruitment_plot_HF_legend`](tables_figures/SI_figures/recruitment_years/Recrutement_years_HarvardForest.png){ width=50% }

\newpage
## `r recruitment_plot_ZOF_legend`

![`r recruitment_plot_ZOF_legend`](tables_figures/SI_figures/recruitment_years/Recrutement_years_Zofin.png){ width=50% }

\newpage
## `r recruitment_plot_NIO_legend`

![`r recruitment_plot_NIO_legend`](tables_figures/SI_figures/recruitment_years/Recrutement_years_Niobara.png){ width=50% }

\newpage
## `r recruitment_plot_LT_legend`

![`r recruitment_plot_LT_legend`](tables_figures/SI_figures/recruitment_years/Recrutement_years_NewMexico.png){ width=50% }

\newpage
## `r recruitment_plot_CB_legend`

![`r recruitment_plot_CB_legend`](tables_figures/SI_figures/recruitment_years/Recrutement_years_CedarBreaks.png){ width=50% }

\newpage
## `r recruitment_plot_SC_legend`

![`r recruitment_plot_SC_legend`](tables_figures/SI_figures/recruitment_years/Recrutement_years_ScottyCreek.png){ width=50% }

@@

![`r decadal_plot_CB_legend`](tables_figures/SI_figures/decadal_plots/CB.png){ width=100% }

\newpage
## `r decadal_plot_SC_legend`
